---
title: "GBP Macro Report"
#subtitle: 
author: Andrew Benito
date: last-modified
format: beamer
execute:
  warning: false
  message: false
  echo: false
---

```{r}
#| include: false
source(here::here("setup", "01_create_data.R"))
source(here::here("setup", "02_plots_ois_gilts_fx_equity.R"))
source(here::here("setup", "07_BoE-forecasts.R"))
source(here::here("setup", "08_mpc_voting.R"))

```

# UK Macro views

1. **Upside inflation risk** resulting from the past pattern of BoE forecast errors and underestimating the UK's supply-side weakness.   

2. **Bond market volatility** amplified by fiscal sustainability concerns. Weak supply-side performance and 'fiscal fatigue' raise questions about fiscal capacity and inflation persistence. In these episodes, yield curve steepening and risk premium episodes should be expected. Fiscal risk premiums tilt the BoE hawkish.   

3. Authorities will mitigate yield curve steepening by active **debt management**, reducing the average maturity of new issuance. HMT can prioritise spending restraint and tax rises with lower implementation risk (eg re-profiling fixed investment plans).   


# GBP Market developments I

```{r}
#| label: fig-ois
#| fig-cap: "OIS and Sterling"
#| results: hide

plot.ois.gbp
```

# GBP Market developments II

```{r}
#| label: fig-gilts
#| fig-cap: "Gilts and Equities"
#| results: hide

plot.gilts.eq
```

# Evolving macro forecasts

```{r}
#| label: fig-infl-successive
#| fig-cap: "Successive BoE Inflation Forecasts"
#| results: hide

infl.forecasts.latest
```

# Successive OIS curves tilting hawkish


```{r}
#| label: fig-ois-successive
#| fig-cap: "Successive OIS Curves"
#| results: hide

ois2
```


# A test of fiscal fatigue?


```{r}
#| label: fig-fatigue
#| fig-cap: "Fiscal fatigue and risk premium in OIS and Sterling"

ois_fatigue
```



# International spillovers in bond markets

```{r}
#| label: fig-rigobon
#| fig-cap: "International spillovers"
#| results: hide

source(here("setup", "04_Rigobon.R"))

plot_structural_shocks

```


# MPC Voting: increased disagreement



```{r}
#| label: hist-votes
#| fig-cap: "A weaker Consensus in MPC rate decisions"

plot.disagreement.t + mpc_agree_plot
```




# UK Macro views

1. **Upside inflation risk** resulting from the past pattern of BoE forecast errors and underestimating the UK's supply-side weakness.   

2. **Bond market volatility** amplified by fiscal sustainability concerns. Weak supply-side performance and 'fiscal fatigue' raise questions about fiscal capacity and inflation persistence. In these episodes, yield curve steepening and risk premium episodes should be expected. Fiscal risk premiums tilt the BoE hawkish.   

3. Authorities will mitigate yield curve steepening by active **debt management**, reducing the average maturity of new issuance. HMT can prioritise spending restraint and tax rises with lower implementation risk (eg re-profiling fixed investment plans).
