---
title: | 
   | GBP Report 
#subtitle: 
author: "Andrew Benito"
date: last-modified
toc: true
format: pdf
execute:
  warning: false
  message: false
  echo: false    
---

```{r}
#| include: false
source(here::here("setup", "01_OIS_curves.R"))
```

# Recent market developments
## Latest developments

```{r}
#| label: fig-2y-daily
#| fig-cap: "2y OIS, changes"
#| out-width: "90%"

plot.daily.60d
```


## Evolving market pricing and OIS
To provide some context for those latest developments, @fig-rates-historical shows how GBP OIS curves have evolved over time, alongside Bank Rate.


```{r}
#| label: fig-rates-historical
#| fig-cap: "Evolution of GBP OIS"
#| out-width: "90%"

ois1
```

@fig-rates-recent zooms in on how OIS curves and Bank Rate have evolved over the past year. 

```{r}
#| label: fig-rates-recent
#| fig-cap: "More Recent Evolution of GBP OIS"
#| out-width: "90%"

ois2
```

## Term spreads: latest developments
2y5y spreads
2y10y spreads
10y30y spreads


# Market reactions and macro news at BoE policy events
## Analysing monetary policy shocks



# Evolving macro forecasts and macro news


# MPC preferences and forecasts
## lambda update

