---
title: | 
   | GBP Rates Report 
#subtitle: 
author: "Andrew Benito"
date: last-modified
toc: true
format: 
  pdf:
    mainfont: "Source Sans Pro"
    sansfont: "Source Sans Pro"
    fontsize: 11pt
    linestretch: 1.4
    include-in-header: 
      text: |
        \usepackage{xcolor}
        \usepackage{sectsty}
        \definecolor{darkblue}{RGB}{0,51,102}        
        \sectionfont{\color{darkblue}}  % sets colour of sections
        \subsectionfont{\color{darkblue}}
execute:
  warning: false
  message: false
  echo: false    
---


```{r}
#| include: false
source(here::here("setup", "01_create_data.R"))
source(here::here("setup", "02_plots_ois_gilts_fx_equity.R"))
source(here::here("setup", "08_mpc_voting.R"))

```

\newpage

# Market developments
## Latest developments: the past three months

Summarising some key developments in Sterling markets in the past 3 months:   
- **OIS rates** have seen a 'round-trip', rallying in June by 25bp, selling-off in July and subsequently reversing those moves. 
- **Sterling** strengthened against the US Dollar in June (while OIS rallied), and then reversed that (as OIS rates sold-off).  
- **Gilt yields** rallied (by less than OIS) in June, and that decline has been more than reversed since then and especially at the long-end. 25y Gilt yields end the period a full 40bp higher.  
- **Equities** rallied strongly by a cumulative 5% over the period, and concentrated in the period from July to mid-August. 

This pattern points to greater fiscal risks entering into market pricing, especially in last August. While fiscal slippage boosts aggregate demand, some crowding-out from these interest rate responses should also be expected. In the past month, fewer rate cuts have been priced to take place over the next year, and this has tended to weigh on Equity prices. Internationally, these interest rate effects appear to have emanated from the US and UK more than from Japan or Germany. 


\newpage

```{r}
#| label: fig-ois
#| fig-cap: "OIS and Sterling"

plot.ois.gbp
```

```{r}
#| label: fig-gilts
#| fig-cap: "Gilts and Equities"

plot.gilts.eq
```

## Spreads (term spreads, swap spreads)

**Term spreads** Setting some historical context for the recent spread of 10y Gilt yields relative to 2y Gilts yields is @fig-2yv10y. This shows how a term spread has been restored relative to different periods over the past 10 years and can distingsuish severa phases over that period.  


```{r}
#| label: fig-2yv10y
#| fig-cap: "Gilt yields"

plot2y_v_10y +
  plot5y_v_25y +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

The 10s25s term spread has widened in recent weeks, and by more than the 2s5s and 5s10s spreads (@fig-spreads).  

```{r}
#| label: fig-spreads
#| fig-cap: "Term spreads in Gilt yields"
#| fig-height: 3

plot2s5s +
  plot5s10s +
  plot10s25s +
  plot_layout(
    axis_titles = "collect",
    ncol = 3,
    heights = 0.7,
    widths = 1,
    guides = "collect"
  ) &
  theme(plot.margin = margin(2, 2, 2, 2)) &
  ylim(range(
    c(
      layer_data(plot2s5s)$y,
      layer_data(plot5s10s)$y,
      layer_data(plot10s25s)$y
    ),
    na.rm = TRUE
  ))
```

**Swap spreads** shed light on potential effects from bond supply in different markets. Swap spreads have moved more deeply negative in recent weeks and by more in the UK than in the US or Germany.  



Fiscal developments have likely played a role in the rise in term spreads and widening of swap spreads. @fig_swsp shows 5y swap spreads (see here for a discussio). 

[contribution from OIS and bond yields to changes in swap spreads in UK, US, GER since 2022-01-01]


```{r}
#| label: fig-swsp
#| fig-cap: "5y swap spreads"

plot.swsp5y + plot_contributions_5yswsp + plot_layout(widths = c(2,3))

```



# Evolving market pricing and OIS curves
Reflecting related developments over a longer period, @fig-ois-historical shows how GBP OIS curves have evolved over time, alongside the MPC's Bank Rate decisions. Research has shown how the persistent forecast errors shown in this chart reflect a gradual process of learning about slow-moving, medium-term factors such as the neutral and trend productivity growth.

@fig-ois-recent zooms-in on how OIS curves and Bank Rate have evolved over the past year, during which the BoE has lowered its policy rate five times. 


```{r}
#| label: fig-ois-historical
#| fig-cap: "Evolution of GBP OIS curves"

ois1
```


```{r}
#| label: fig-ois-recent
#| fig-cap: "More Recent Evolution of GBP OIS curves"

ois2
```

**Neutral rates** We limit our discussion of neutral rates. @fig-neutral shows estimates since the pandemic based on a survey of market participants in Sterling markets. In principle, this survey should embody informed assessments of the impact on neutral rates from a broad set of influences. I simulate individual responses based on the reported summary statistics published by the BoE in its Market Participants Survey (MaPS).

```{r}
#| label: fig-neutral
#| fig-cap: "Market Participants Survey: Neutral rate estimates"

source(here::here("setup", "03_neutral_r.R"))
boxplot.neutral
```




# International spillovers in sovereign bond markets

What of international spillovers in bond markets? Using the Rigobon (2003) methodology, we can estimate the extent to which international spillovers have contributed to fluctuations in 10y Gilt yields. With bond markets 'spilling over' to one another, this aims to quantify to what extent different markets were 'in the driving seat' at a point in time.   

These estimates point to the US and UK being the main 'drivers' of international bond markets over the past decade, with Japan and Germany have been on the receiving end of these spillovers. 


```{r}
#| label: fig-rigobon
#| fig-cap: "International spillovers"
#| results: hide

source(here("setup", "04_Rigobon.R"))

plot_structural_shocks

```

# Fiscal risks
## The (international) risk of fiscal fatigue

One Fiscal fatigue occurs where a further fiscal effort is needed to stabilise public debt-to-GDP, but the political will to implement the necessary tax rises or restraint in public spending wanes. 

We estimate the UK still has a required fiscal effort of 3% of GDP to stabilise its Debt-to-GDP ratio. While completing that required fiscal effort is forecast by the OBR, the issue is whether that political will exists, absent risk premiums being priced by markets. That risk of fiscal fatigue likely lies behind the periodic risk premiums in Sterling markets, including reflected in term spreads and swap spreads shown previously.

Uncertainty in this estimate of fiscal effort means it is worth showing the debt-stabilising primary balance for different combinations of 'r' (effective nominal interest rate on government debt) and 'g' (expected nominal growth rate). Estimates in @tab-dspb suggest the debt-stabilising primary balance would be aorund 0.5-1.0% of GDP, compared with a primary balance at -2.0% in 2024.  

By international standards, this is xxx. [France stands out.]

Populism likely adds to this risk of fiscal fatigue, including in the UK and France. 


## Macro and fiscal news at UK fiscal events

Fig for indiv fiscal events


A fiscal downgrade of 30bn would, based on this relationship, point to tightening of 
A key issue in the Budget will be how front-loaded is that tightening. 


```{r}
#| label: tab-fiscal-dspb
#| fig-cap: "Debt-stabilising primary balance"
#| results: hide

source(here::here("setup", "05_fiscal_effort.R"))
print(gt.dat) 
```


```{r}
#| label: fig-fiscal-fm
#| fig-cap: "Estimates of Required Fiscal Effort"
#| results: hide

plot_effort
```


# Evolving macro forecasts and macro news

Those market reactions may owe partly to the MPC's own assessment of the UK macro outlook, eg. under the so-called information channel of monetary policy. 

**Medium-term** Over the past decade or more, the MPC has tended to revise higher its forecasts for inflation, revise its outlook for unemployment lower, without an obvious bias in its forecasts for GDP growth. This points points to a pattern of the BoE having had to respond to negative supply shocks, updating its assessments in the light of that repeated pattern. 

The weakness of productivity growth, negative terms of trade shocks (e.g. Brexit, pandemic, Ukraine), and the impact of these on real incomes, have all contributed to this pattern of revisions and evolving macro forecasts.

**Recent updates** Looking over the past six MPC forecasts...

```{r}
#| label: plot-fe 
#| fig-cap: "Distribution of BoE Forecast errors"

source(here::here("setup", "07_BoE-forecasts.R"))

p3 + p4 + p5

```

## Evolving inflation forecasts
Based on the pattern of past BoE forecast errors, a realistic path to overshooting the 2% inflation target, is that the MPC produces too optimistic inflation forecasts and retaining too little policy restriction to complete the disinflation. 



```{r}
#| label: plot-infl-latest 
#| fig-cap: "Successive BoE Inflation Forecasts"


infl.forecasts.latest

```



## Risk premium episodes tilt hawkish for monetary policy

Early-July was (anoteher) case study of the fiscal fatigue risk premium. The UK Parliament failed to pass a welfare bill of welfare restraint (modifying the bill before without fiscal savings before passing it). The Chancellor's tears at the despatch box were widely reported, alongside speculation of a risk Reeves would be replaced and with that the current fiscal rules revised. 

The hawkish implied response was reflected in 2y and 5y OIS rates rising by 10-15bp over the following week (@fig-fatigue). THat was associated with a 2% decline in GBPUSD. 

```{r}
#| label: fig-fatigue
#| fig-cap: "Fiscal fatigue and risk premium in OIS and Sterling"

ois_fatigue
```


# MPC Voting: increased disagreement 

We inspect MPC voting patterns to illustrate two points. First, the rising extent of disagreement among MPC members. Second, how current MPC member voting compares with that of past members.   

We make the further point that the rising disagreement among MPC members reflects the increasing role for supply-shocks, and more complex macroeconomic environment, in which MPC members are making decisions.


```{r}
#| label: hist-votes
#| fig-cap: "A weaker Consensus in MPC rate decisions"

plot.disagreement.t + hist.votes
```

Rising disagreement among MPC members culminated in the August 2025 MPC vote requiring a second vote among Committee members in order to secure a clear majority for the rate cut decision.


```{r}
#| label: member-patterns
#| fig-cap: "MPC disagreement and preferences"
#| out-width: "100%"

plot.member_patterns
```


Increasingly nuanced macroeconomic assessments are producing more closely contested votes and heightened disagreement among MPC members. This development places greater emphasis on the quality of macroeconomic analysis and forecasts that underpin MPC decisions. While such complex judgements reinforce the importance of maintaining independence from political pressures, the forecast errors documented earlier pose a risk to both the credibility of the MPC's analytical process and public confidence in its institutional independence. This may leave Sterling markets more sensitive to international spillovers and to political and fiscal news. 


```{r}
#| label: plot-correls-gbp
#| fig-cap: "Rolling correlation: 10y yields and GBPUSD"
#| results: hide

source(here::here("setup", "09_bouts_risk_premium.R"))

correl.plot.10y.gbp 
```





```{r}
#| label: plot-correls-ftse
#| fig-cap: "Rolling correlation: 10y yields and Equities"
#| results: hide

correl.plot.10y.ftse 
```



# Conclusions: Macro views

1. Government bonds provide less effective portfolio hedging against equity weakness when negative supply shocks become more frequent. In supply-shock driven downturns, both equities and bonds can decline simultaneously, reducing the traditional negative correlation that makes bonds effective diversifiers.
2. Historical forecast error patterns suggest the MPC faces material upside inflation risks if it maintains current policy settings. The Bank's tendency to overestimate the pace of disinflation, combined with insufficient recognition of the UK's structural supply-side constraints, creates conditions where inflation could persistently exceed the 2% target despite current monetary restriction.
3. Fiscal sustainability concerns amplify bond market volatility, particularly during data-sensitive periods when investors reassess inflation trajectories. Disappointing economic data, coming from weak supply-side performance, simultaneously raise questions about fiscal capacity and inflation persistence. In these episodes, we should expect yield curve steepening. 
4. The authorities can mitigate yield curve steepening through active debt maturity management, specifically by reducing the average maturity of new issuance. HMT should prioritise other forms of spending restraint and taxes that face lower implementation risk. Front-loading offers similar benefits. 
5. If the authorities instead prioritise back-loading fiscal adjustment and/or with higher implementation risk, this would contribute to further steepening.  
6. Fiscal risk premium episodes tilt hawkish for the BoE. Our sense is that this view surprises most analysts. 
 




Fisal fatigue