---
title: | 
   | GBP Macro Report 
#subtitle: 
author: "Andrew Benito"
date: last-modified
toc: true
format: 
  pdf:
    mainfont: "Source Sans Pro"
    sansfont: "Source Sans Pro"
    fontsize: 11pt
    linestretch: 1.4
    include-in-header: 
      text: |
        \usepackage{xcolor}
        \usepackage{sectsty}
        \definecolor{darkblue}{RGB}{0,51,102}        
        \sectionfont{\color{darkblue}}  % sets colour of sections
        \subsectionfont{\color{darkblue}}
execute:
  warning: false
  message: false
  echo: false    
---


```{r}
#| include: false
source(here::here("setup", "01_create_data.R"))
source(here::here("setup", "02_plots_ois_gilts_fx_equity.R"))
source(here::here("setup", "08_mpc_voting.R"))

```

\newpage

# Market developments
## Latest developments: the past three months

Summarising some key developments in Sterling markets in the past 3 months:   
- **OIS rates** have seen a 'round-trip', rallying in June by 25bp, selling-off in July and August in moves subsequently reversed.   
- **Sterling** strengthened against the US Dollar in June (while OIS rallied), and then reversed that (as OIS rates sold-off).  
- **Gilt yields** rallied (by less than OIS) in June, and that decline has been more than reversed since then and especially at the long-end. Sharp rises in August in 25y+ Gilt yields reversed into September.  
- **Equities** rallied strongly by a cumulative 5% over the period, and concentrated in the period from July to mid-August. 

This pattern points to greater fiscal and inflation risks entering into market pricing, before subsiding. While fiscal slippage would boost aggregate demand, some crowding-out from the interest rate responses would offset that. During the past month, fewer rate cuts have been priced to take place over the next year (despite more cuts being priced into Fed policy). Internationally, interest rate effects appear to have emanated from the US and UK more than from Japan or Germany. 


\newpage

```{r}
#| label: fig-ois
#| fig-cap: "OIS and Sterling"

plot.ois.gbp
```

```{r}
#| label: fig-gilts
#| fig-cap: "Gilts and Equities"

plot.gilts.eq
```

# Evolving OIS curves
@fig-ois-historical shows how GBP OIS curves have evolved over time, alongside the MPC's Bank Rate decisions. Persistent forecast errors shown in this chart reflect a gradual process of markets learning about slow-moving, medium-term factors such as the neutral interest rate and trend productivity growth. 

@fig-ois-recent zooms-in on how OIS curves and Bank Rate have evolved over the past year, during which the BoE has lowered its policy rate five times to 4.0% in its August MPC meeting. 


```{r}
#| label: fig-ois-historical
#| fig-cap: "Evolution of GBP OIS curves"

ois1
```


```{r}
#| label: fig-ois-recent
#| fig-cap: "More Recent Evolution of GBP OIS curves"

ois2
```

**Neutral rates** We limit our discussion of neutral rates. @fig-neutral shows estimates since the pandemic based on a survey of market participants in Sterling markets. In principle, this survey should embody informed assessments of the impact on neutral rates from a broad set of influences. I simulate individual responses based on the reported summary statistics published by the BoE in its Market Participants Survey (MaPS).

```{r}
#| label: fig-neutral
#| fig-cap: "Market Participants Survey: Neutral rate estimates"

source(here::here("setup", "03_neutral_r.R"))
boxplot.neutral
```

# International spillovers in sovereign bond markets

What of international spillovers in bond markets? Using the Rigobon (2003) methodology, we can estimate the extent to which international spillovers have contributed to fluctuations in 10y Gilt yields and other 10y yields. With bond markets 'spilling over' to one another, this aims to quantify to what extent different markets were 'in the driving seat' at a point in time.   

These estimates point to the US and UK being the main 'drivers' of international bond markets over the past decade, with Japan and Germany have been on the receiving end of these spillovers. 


```{r}
#| label: fig-rigobon
#| fig-cap: "International spillovers"
#| results: hide

source(here("setup", "04_Rigobon.R"))

plot_structural_shocks

```

# Fiscal risks
## The (international) risk of fiscal fatigue

'Fiscal fatigue' occurs when a further fiscal effort is needed to stabilise public debt-to-GDP, but the political will to implement the necessary tax rises or restraint in public spending wanes. Fiscal fatigue is a key risk for the UK and other countries with high Debt-to-GDP ratios and challenges from populism. 

We estimate that the UK still has a required fiscal effort of 3% of GDP to stabilise its Debt-to-GDP ratio. While completing that required fiscal effort is forecast by the OBR on current fiscal plans, the issue including for markets is whether the political will exists, absent risk premiums being priced by markets. 

The UK's failure to pass a Welfare bill with welfare savings in early-July was a clear instance of fiscal fatigue, associated with speculation about replacing Chancellor Reeves and loosening the current fiscal rules (see below). The risk of fiscal fatigue implies periodic risk premiums in Sterling markets, including reflected in term spreads and swap spreads shown previously.

The required fiscal effort depends on 'r' (effective nominal interest rate on government debt) and 'g' (expected nominal growth rate). Estimates in @tab-dspb suggest the debt-stabilising primary balance would be aorund 0.5-1.0% of GDP, compared with a primary balance at -2.0% in 2024.  

By international standards, this is at the upper end of required fiscal effort, leaving the UK prone to periodic bouts of risk premiums, especially associated with political risks. France provides another example of a large required fiscal effort.


## Macro news at UK fiscal events

Fiscal tightening or loosening follows results from the amount of macro and fiscal news at a fiscal event (see @fig-fiscal-fm). On these estimates 




A fiscal downgrade of 30bn would, based on this relationship, point to tightening of 
A key issue in the Budget will be how front-loaded is that tightening. 


```{r}
#| label: tab-fiscal-dspb
#| fig-cap: "Debt-stabilising primary balance"
#| results: hide

source(here::here("setup", "05_fiscal_effort.R"))
gt.dat 
```


```{r}
#| label: fig-fiscal-fm
#| fig-cap: "Estimates of Actual and Required Fiscal Effort"
#| results: hide

source(here::here("setup", "06_fiscal_choices.R"))
responses.scatter + plot_effort
```


# Evolving macro forecasts and macro news

Those market reactions may owe partly to the MPC's own assessment of the UK macro outlook, eg. under the so-called information channel of monetary policy. 

**Medium-term** Over the past decade or more, the MPC has tended to revise higher its forecasts for inflation, revise its outlook for unemployment lower, without an obvious bias in its forecasts for GDP growth. This points points to a pattern of the BoE having had to respond to negative supply shocks, updating its assessments in the light of that repeated pattern. 

The weakness of productivity growth, negative terms of trade shocks (e.g. Brexit, pandemic, Ukraine), and the impact of these on real incomes, have all contributed to this pattern of revisions and evolving macro forecasts.

**Recent updates** Looking over the past six MPC forecasts...

```{r}
#| label: plot-fe 
#| fig-cap: "Distribution of BoE Forecast errors"

source(here::here("setup", "07_BoE-forecasts.R"))

p3 + p4 + p5

```

## Evolving inflation forecasts
Based on the pattern of past BoE forecast errors, a realistic path to overshooting the 2% inflation target, is that the MPC produces too optimistic inflation forecasts and retaining too little policy restriction to complete the disinflation. 



```{r}
#| label: plot-infl-latest 
#| fig-cap: "Successive BoE Inflation Forecasts"


infl.forecasts.latest

```



## (Fiscal) Risk premium episodes tilt hawkish for monetary policy

Early-July was (anoteher) case study of the fiscal fatigue risk premium. The UK Parliament failed to pass a welfare bill of welfare restraint (modifying the bill before without fiscal savings before passing it). The Chancellor's tears at the despatch box were widely reported, alongside speculation of a risk Reeves would be replaced and with that the current fiscal rules revised. 

The hawkish implied response was reflected in 2y and 5y OIS rates rising by 10-15bp over the following week (@fig-fatigue). THat was associated with a 2% decline in GBPUSD. 

```{r}
#| label: fig-fatigue
#| fig-cap: "Fiscal fatigue and risk premium in OIS and Sterling"

ois_fatigue
```


# MPC Voting: increased disagreement 

We inspect MPC voting patterns to illustrate two points. First, the rising extent of disagreement among MPC members. Second, how current MPC member voting compares with that of past members.   

We make the further point that the rising disagreement among MPC members reflects the increasing role for supply-shocks, and more complex macroeconomic environment, in which MPC members are making decisions.


```{r}
#| label: hist-votes
#| fig-cap: "A weaker Consensus in MPC rate decisions"

plot.disagreement.t + mpc_agree_plot
```

Rising disagreement among MPC members culminated in the August 2025 MPC vote requiring a second vote among Committee members in order to secure a clear majority for the rate cut decision.


```{r}
#| label: member-patterns
#| fig-cap: "MPC disagreement and preferences"
#| out-width: "100%"

plot.member_patterns
```


Increasingly nuanced macroeconomic assessments are producing more closely contested votes and heightened disagreement among MPC members. This development places greater emphasis on the quality of macroeconomic analysis and forecasts that underpin MPC decisions. While such complex judgements reinforce the importance of maintaining independence from political pressures, the forecast errors documented earlier pose a risk to both the credibility of the MPC's analytical process and public confidence in its institutional independence. This may leave Sterling markets more sensitive to international spillovers and to political and fiscal news. 


```{r}
#| label: plot-rp 
#| fig-cap: "Correlation UK/US 10y Yield Gap and GBPUSD"

source(here::here("setup", "09_bouts_risk_premium.R"))

correl.plot.10y.gbp

```


# Sell-side forecasts for Bank Rate


```{r}
#| label: plot-sellside 
#| fig-cap: "Sell-side Forecasts for BoE Bank Rate"

source(here::here("setup", "10_sell-side.R"))

plot.sellside.violin

```




# Conclusions: Macro views


1. UK macro faces a higher-than-normal incidence of negative supply shocks. This generates a 'stagflationary' impulse and complicates the MPC's task of achieving the 2.0% inflation target. In this environment, government bonds provide a less effective hedge which would structurally weaken demand for bonds.  
2. The pattern of past forecast errors points to a material upside inflation risk, if that pattern is repeated partly on underestimating the UK's supply-side weakness.  
3. Fiscal sustainability concerns amplify bond market volatility, particularly at the long-end and during data-driven periods. Disappointing economic data, coming from weak supply-side performance, simultaneously raise questions about fiscal capacity and inflation persistence. In these episodes, yield curve steepening and risk premium episodes should be expected.  
4. The authorities can mitigate yield curve steepening through active debt management, now a bigger feature of the UK policy mix, reducing the average maturity of new issuance. A second policy option to contain steepening is that HMT prioritises restraint in public spending and tax rises that face lower implementation risk (eg re-profiling fixed investment plans). Back-loading fiscal adjustment has the opposite effect.  
5. Fiscal fatigue is therefore a key form of fiscal risk that bears watching. This risk would crystallise if the authorities instead prioritise back-loading fiscal adjustment and/or with higher implementation risk, this would contribute to further steepening.  
6. Fiscal risk premium episodes tilt hawkish for the BoE. Our sense is that this view surprises most analysts. Our current forecast is for the BoE to cut Bank Rate to 3.75% in November, but that forecast is subject to a high degree of uncertainty.
 