---
title: | 
   | GBP Rates Report 
#subtitle: 
author: "Andrew Benito"
date: last-modified
toc: true
format: 
  pdf:
    mainfont: "Source Sans Pro"
    sansfont: "Source Sans Pro"
    fontsize: 11pt
    linestretch: 1.4
    include-in-header: 
      text: |
        \usepackage{xcolor}
        \usepackage{sectsty}
        \definecolor{darkblue}{RGB}{0,51,102}        
        \sectionfont{\color{darkblue}}  % sets colour of sections
        \subsectionfont{\color{darkblue}}
execute:
  warning: false
  message: false
  echo: false    
---


```{r}
#| include: false
source(here::here("setup", "01_create_data.R"))
source(here::here("setup", "02_plots_ois_gilts_fx_equity.R"))
source(here::here("setup", "08_mpc_voting.R"))

```

\newpage

# Market developments
## Latest developments: the past three months

Summarising some key developments in Sterling markets in the past 3 months:   
- **OIS rates** have seen a 'round-trip', rallying in June by 25bp and subsequently reversing that move. 
- **Sterling** strengthened against the US Dollar in June (while OIS rallied), and then reversed that (as OIS rates sold-off).  
- **Gilt yields** rallied (by less than OIS) in June, and that decline has been more than reversed since then and especially at the long-end. 25y Gilt yields end the period a full 40bp higher.  
- **Equities** rallied strongly by a cumulative 5% over the period, and concentrated in the period from July to mid-August. 



```{r}
#| label: fig-ois
#| fig-cap: "OIS and Sterling"

plot.ois.gbp
```

```{r}
#| label: fig-gilts
#| fig-cap: "Gilts and Equities"

plot.gilts.eq
```

## Spreads

**Term spreads** Setting some historical context for the recent spread of 10y Gilt yields relative to 2y Gilts yields is @fig-2yv10y. This shows how a term spread has been restored relative to different periods over the past 10 years and can distingsuish severa phases over that period.  


```{r}
#| label: fig-2yv10y
#| fig-cap: "Gilt yields"

plot2y_v_10y +
  plot5y_v_25y +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

The 10s25s term spread has widened in recent weeks, and by more than the 2s5s and 5s10s spreads (@fig-spreads).  

```{r}
#| label: fig-spreads
#| fig-cap: "Term spreads in Gilt yields"
#| fig-height: 3

plot2s5s +
  plot5s10s +
  plot10s25s +
  plot_layout(
    axis_titles = "collect",
    ncol = 3,
    heights = 0.7,
    widths = 1,
    guides = "collect"
  ) &
  theme(plot.margin = margin(2, 2, 2, 2)) &
  ylim(range(
    c(
      layer_data(plot2s5s)$y,
      layer_data(plot5s10s)$y,
      layer_data(plot10s25s)$y
    ),
    na.rm = TRUE
  ))
```

**Swap spreads** help shed light on potential effects from bond supply in different markets. Swap spreads have moved more deeply negative in recent weeks and by more in the UK than in the US or Germany.  

[some detail]

Fiscal developments have likely played a role in the rise in term spreads and widening of swap spreads (see here for a discussio). 

[contribution from OIS and bond yields to changes in swap spreads in UK, US, GER since 2022-01-01]


# Evolving market pricing and OIS curves
Reflecting related developments over a longer period, @fig-ois-historical shows how GBP OIS curves have evolved over time, alongside the MPC's Bank Rate decisions. Research has shown how the persistent forecast errors shown in this chart reflect a gradual process of learning about slow-moving, medium-term factors such as the neutral and trend productivity growth.

@fig-ois-recent zooms-in on how OIS curves and Bank Rate have evolved over the past year.   [check this chart]


```{r}
#| label: fig-ois-historical
#| fig-cap: "Evolution of GBP OIS curves"

ois1
```


```{r}
#| label: fig-ois-recent
#| fig-cap: "More Recent Evolution of GBP OIS curves"

ois2
```

**Neutral rates** We limit our discussion of neutral rates. @fig-neutral shows estimates since the pandemic based on a survey of market participants in Sterling markets. In principle, this survey should embody informed assessments of the impact on neutral rates from a broad set of influences. I simulate individual responses based on the reported summary statistics published by the BoE in its Market Participants Survey (MaPS).

```{r}
#| label: fig-neutral
#| fig-cap: "Market Participants Survey: Neutral rate estimates"

source(here::here("setup", "03_neutral_r.R"))
boxplot.neutral
```




# International spillovers in sovereign bond markets

What of international spillovers? Using the Rigobon (2003) methodology, we can estimate the extent to which international spillovers have contributed to fluctuations in 10y Gilt yields.  




```{r}
#| label: fig-rigobon
#| fig-cap: "International spillovers"
#| results: hide

source(here("setup", "04_Rigobon.R"))

plot_structural_shocks

```

# Fiscal risks
## The (international) risk of fiscal fatigue

Required fiscal effort and how UK compares.

## Macro and fiscal news at UK fiscal events

A fiscal downgrade of 30bn would, based on this relationship, point to tightening of 
A key issue in the Budget will be how front-loaded is that tightening. 


# Market reactions and macro news at recent BoE policy events
## Analysing monetary policy shocks

We now review how financial markets -- OIS, Gilts, Foreign exchange and Equities -- reacted to news at BoE policy events. These reactions are summarised in @fig-mon-factors. 



```{r}
#| label: fig-mon-factors
#| fig-cap: "Asset price reactions at MPC announcements"
#| fig-height: 6

source(here::here("setup", "06_mon_pol_shocks.R"))
plot.reactions
```



# Evolving macro forecasts and macro news

How has the MPC's own assessment of the UK macro outlook evolved? 

**Medium-term** Over the past decade or more, the MPC has tended to revise higher its forecasts for inflation, revise its outlook for unemployment lower, without an obvious bias in its forecasts for GDP growth. This points points to a pattern of the BoE having had to respond to negative supply shocks, updating its assessments in the light of that repeated pattern. 

The weakness of productivity growth, negative terms of trade shocks (e.g. Brexit, pandemic, Ukraine), and the impact of these on real incomes, have all contributed to this pattern of revisions and evolving macro forecasts.

**Recent updates** Looking over the past six MPC forecasts...

```{r}
#| label: plot-fe 
#| fig-cap: "Distribution of BoE Forecast errors"

source(here::here("setup", "07_BoE-forecasts.R"))

p3 + p4 + p5

```

# MPC Voting: increased disagreement 

We inspect MPC voting patterns to illustrate two points. First, the rising extent of disagreement among MPC members. Second, how current MPC member voting compares with that of past members.   

```{r}
#| label: hist-votes
#| fig-cap: "A weaker Consensus in MPC rate decisions"

plot.disagreement.t + hist.votes
```

Rising disagreement among MPC members culminated in the August 2025 MPC vote requiring a second vote among Committee members in order to secure a clear majority for the rate cut decision.


```{r}
#| label: member-patterns
#| fig-cap: "MPC disagreement and preferences"
#| out-width: "100%"

plot.member_patterns
```


Increasingly nuanced macroeconomic assessments are producing more closely contested votes and heightened disagreement among MPC members. This development places greater emphasis on the quality of macroeconomic analysis and forecasts that underpin MPC decisions. While such complex judgements reinforce the importance of maintaining independence from political pressures, the forecast errors documented earlier pose a risk to both the credibility of the MPC's analytical process and public confidence in its institutional independence. This may leave Sterling markets more sensitive to international spillovers and to political and fiscal news. 

# Views

Based on the pattern of past BoE forecast errors, a realistic path to overshooting the 2% inflation target, is that the MPC produces too optimistic inflation forecasts and retaining too little policy restriction to complete the disinflation. 
