---
title: | 
   | GBP Rates Report 
#subtitle: 
author: "Andrew Benito"
date: last-modified
toc: true
format: pdf
execute:
  warning: false
  message: false
  echo: false    
---

```{r}
#| include: false
source(here::here("setup", "01_create_data.R"))
source(here::here("setup", "02_plots_ois_gilts.R"))
#source(here::here("setup", "04_Rigobon.R"))

```

# Recent market developments
## Latest developments

@fig-2y-daily shows the (cumulative) changes in 2y, 5y and 10y OIS rates over the past two months. A sell-off in rates has been more pronounced among 10year rates than 2year rates. It also coincided 

```{r}
#| label: fig-2y-daily
#| fig-cap: "2y OIS, changes"
#| out-width: "90%"

plot.cumul.60d
```

## Spreads

**Term spreads** Setting some context for the recent spread of 10y Gilt yields relative to 2y Gilts yields is @fig-2yv10y. This shows how a term spread has been restored relative to different periods over the past 10 years.  


```{r}
#| label: fig-2yv10y
#| fig-cap: "Gilt yields: 2y v 10y"
#| out-width: "90%"

plot2y_v_10y
```

The 10s25s term spread has widened in recent weeks, and by more than the 2s5s and 5s10s spreads (@fig-spreads).  

```{r}
#| label: fig-spreads
#| fig-cap: "Term spreads in Gilt yields"
#| out-width: "110%"

plot2s5s + plot5s10s + plot10s25s
```

**Swap spreads** have also widened in recent weeks, as shown in @fig-swap-spreads. This reflects both the rise in Gilt yields, and a fall in swap rates. Swap spreads have been volatile over the past decade or more, reflecting changes in market liquidity, risk premia, and other factors.

Fiscal developments have likely played a role in the rise in term spreads and widening of swap spreads. 

**Neutral rates** We limit our discussion of neutral rates. @fig-neutral shows estimates based on a survey of market participants. I simulate individual responses based on the reported summary statistics published by the BoE in its Market Participants Survey.  



# Evolving market pricing and OIS curves
Providing some context for those latest developments, @fig-rates-historical shows how GBP OIS curves have evolved over time, alongside the MPC's Bank Rate decisions.


```{r}
#| label: fig-ois-historical
#| fig-cap: "Evolution of GBP OIS curves"
#| out-width: "90%"

ois1
```

@fig-rates-recent zooms-in on how OIS curves and Bank Rate have evolved over the past year. 

```{r}
#| label: fig-ois-recent
#| fig-cap: "More Recent Evolution of GBP OIS curves"
#| out-width: "90%"

ois2
```


# International spillovers

What of international spillovers? Using the Rigobon (2003) methodology, we can estimate the extent to which international spillovers have contributed to the recent sell-off in GBP 10y Gilt yields.  


```{r}
#| label: fig-rigobon
#| fig-cap: "Term spreads in Gilt yields"
#| out-width: "90%"

#plot.rigobon
```


# Market reactions and macro news at BoE policy events
## Analysing monetary policy shocks

In this section, we review market reactions to BoE policy events, and how these reactions relate to macro news digested by markets at those BoE policy events. This provides further context to recent market and macro developments.



#```{r}
#| include: false
source(here::here("setup", "03_mon_pol_shocks.R"))
#```



# Evolving macro forecasts and macro news

How has the MPC's own assessment of the macro outlook evolved? 

**Medium-term** Over the past decade or more, the MPC has tended to revise higher its forecasts for inflation, revise its outlook for unemployment lower, without an obvious bias in its forecasts for GDP growth. This points points to a pattern of the BoE having had to respond to negative supply shocks, updating its assessments in the light of that repeated pattern. 

The weakness of productivity growth, negative terms of trade shocks (e.g. Brexit, Covid, Ukraine), and the impact of these on real incomes, have all contributed to this pattern of revisions and evolving macro forecasts.

**Recent updates** Looking over the past six MPC forecasts...



# MPC Voting 

We inspect MPC voting patterns to assess two points. First, the current extent of disagreement among MPC members versus history. Second, MPC voting among the current MPC members. 


